---
title: "Project - Analysis of housing price of Korea"
author: "Jae Hyun Lee, jl914"
date: "3/24/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
library(lubridate)
library(grid)
library(gridExtra)
set.seed(642)
```

```{r,include=FALSE}
apartment <- read_csv("gangnam.csv")
apartment <- apartment %>% 
  mutate_if(is.character,as.factor) %>% 
  dplyr::select(-city, -name, -road_name, -district) %>% 
  mutate(price = price/100, date = as_date(str_c(month,"01")), age = year(date)-built) %>% 
  mutate(lprice = log(price), lsize = log(size), lfloor = log(floor), lage = log(age+1))
idx <- apply(apartment,1,function(x){sum(is.na(x))})#is.na(apartment$lage)
apartment <- apartment[idx==0,]
test <- read_csv("test.csv")
test <- test %>% 
  mutate_if(is.character,as.factor) %>% 
  dplyr::select(-city, -name, -road_name, -district) %>% 
  mutate(price = price/100, date = as_date(str_c(month,"01")), age = year(date)-built) %>% 
  mutate(lprice = log(price), lsize = log(area), lfloor = log(floor), lage = log(age+1))
idx <- is.na(test$lfloor)
test <- test[!idx,]
```

## EDA

```{r}
m_price <- apartment %>% 
  group_by(date) %>% 
  dplyr::summarise(lprice = mean(lprice), count = dplyr::n())
ggplot() +
  geom_line(data = m_price, mapping = aes(x = date, y = lprice)) +
  labs(title = "average log(price) of apartment in Korea currency")
```

```{r, include=FALSE}
data <- apartment %>% 
  dplyr::select(lprice,lsize,lage,date)
predictors <- c("lsize", "lage")
T <- nrow(m_price)
ti <- sort(unique(data$date))
betas <- matrix(rep(NA,T*3),ncol = 3)
sigmas <- matrix(rep(NA,T))
n = matrix(rep(NA,T))
resid <- list()
cor_mat <- list()
for(i in 1:T){
  idx <- data$date == ti[i] 
  newdata <- data[idx,]
  model <- lm(data = newdata, lprice~lsize+lage)
  betas[i,] <- model$coefficients
  sigmas[i] <- summary(model)$sigma^2
  n[i] <- nrow(newdata)
  resid[[i]] <- summary(model)$residuals
}
betas <- as.data.frame(betas) %>% 
  `colnames<-`(c(paste0("beta",1:3)))
plot_list <- list()
plot_list[[1]] <- ggplot(data = betas) +
  geom_line(mapping = aes(x = ti, y = beta1)) +
  labs(title = "coefficient of intercept")
plot_list[[2]] <- ggplot(data = betas) +
  geom_line(mapping = aes(x = ti, y = beta2)) +
  labs(title = "coefficient of log(Size)")
#plot_list[[3]] <- ggplot(data = betas) +
#  geom_line(mapping = aes(x = ti, y = beta3)) +
#  labs(title = "coefficient of log(Floor)")
plot_list[[3]] <- ggplot(data = betas) +
  geom_line(mapping = aes(x = ti, y = beta3)) +
  labs(title = "coefficient of log(Age)")
plot_sigma <- ggplot()+
  geom_line(mapping = aes(x = ti, y = sigmas))+
  labs(title = "time varying sigma")
sam <- sample(1:168,2)
```

For the target variable, I decided to log transform original target variable to satisfy normality assumption of model. As a result, we can see that normality assumptions for each time point is not severely violated.

```{r, fig.align='center', fig.height=3.5, fig.width=7}
par(mfrow = c(1,2))
qqnorm(resid[[sam[1]]])
  qqline(resid[[sam[1]]],col = "red")
qqnorm(resid[[sam[2]]]) 
  qqline(resid[[sam[2]]],col = "red")
```

Evolution matrix $G_t$ is not determined yet, but first attempt would be AR(1) model. It will be extended to AR(p) or TAVR(p). Since linear regression assumes that errors are indepedent, I will apply that assumption in my model too. In evolving and updating process, I will use result from course materials and P&W.

I have confirmed that my first inspiration time varying slope is reasonable hen we investigate $\beta$s of each time point. At these plots, we can find that we should fit time varying coefficients rather than constants.

```{r, fig.align='center', fig.height=4, fig.width=7}
grid.arrange(grobs = plot_list,nrow = 2)
```

we can also find that estimated variance of $y_t$ becomes very unstable, thus I plan to use volatile model and large discount factor for $V(v_t) = \nu_t$ to reflect this phenomenon. 

```{r,fig.align='center', fig.height=2, fig.width=4}
plot_sigma
```

## Time varying regression

```{r}
G <- diag(rep(1,3),nrow = 3)
T <- length(ti)
q <- length(betas)
#priors (beta,vt|D0)
nu0 = 100; S0 = 0.3^2
m0 = c(1,1,0)
C0 = diag(diag(cov(betas)),nrow = 3)
#discount factor
b = 0.80; d = 0.95
sm = sa = matrix(rep(0,q*T),nrow = q); sC = sR = array(rep(0,q*q*T),dim = c(q,q,T)) 
sf = sQ = list()
sS = snu = rep(0,T)
nu <- nu0; S = S0; m <- m0; C <- C0
#Foreward filtering
for(t in 1:T){
  idx = data$date == ti[t]
  yt = data$lprice[idx]; ot = length(yt)
  X = as.matrix(cbind.data.frame(intercept = rep(1,sum(idx)),data[idx,-c(1,4)]))    #regression matrix
  
  #evolve
  nu <- b*nu
  sa[,t] <- a <- G%*%m
  sR[,,t] <- R <- G%*%C%*%t(G)/d

  #forecast
  sf[[t]] <- f <- X%*%a
  sQ[[t]] <- Q <- X%*%R%*%t(X) + diag(rep(S,ot),nrow = ot)
  
  #update
  Tt <- solve(t(X)%*%X + c(S)*solve(R))
  sm[,t] <- m <- Tt%*%(t(X)%*%yt + c(S)*solve(R)%*%a)
  sS[t] <- S <- (nu*c(S) + t(yt)%*%yt + c(S)*t(a)%*%solve(R)%*%a - t(m)%*%solve(Tt)%*%m)/(nu +ot)
  snu[t] <- nu <- nu + ot
  sC[,,t] <- C <- Tt/c(S)
}
```


## VAR(1) model for betas

$G_t = G= \begin{bmatrix}\phi_{11}&\cdots&\phi_{14} \\\phi_{21} &\cdots &\phi_{24}\\\phi_{31} &\cdots &\phi_{34} \\ \phi_{41} &\cdots &\phi_{44}\end{bmatrix}$

```{r}
VAR <- function(x,p){
  x <- apply(x,2,function(x){x - mean(x)})
  n <- dim(x)[1] #number of time point
  q <- dim(x)[2] #number of series
  b <- matrix(rep(NA,q*q*p),nrow = q)
  X <- matrix(rep(NA,(n-p)),nrow = n-p)
  for(j in 0:(p-1)){
      X <- cbind(X,x[(p-j):(n-1-j),])
  }
  X <- X[,-1]
  B <- t(X) %*% X  
  for(i in 1:q){
    y <- x[-(1:p),i]
    b[i,] <- solve(B) %*% t(X) %*% y
  }
  return(list(b = b, B = B))
}
```

## R var function

```{r}
library(vars)
betas2 <- ts(beta)
vars::VAR(y = betas2, p = 1,type = "none")
VAR(beta,1)
```

## stationarity check for VAR(1)

```{r}
G <- VAR(betas, 2)$b
#eye4 = diag(rep(1,4),nrow = 4)
#eigen(eye4 - G)
eigen(G)
```

```{r}
inv <- function(m){
  result <- eigen(m)
  vec <- result$vectors
  lamb <- result$values
  return(vec%*%diag(1/lamb,nrow = nrow(vec))%*%solve(vec))
} 
G <- VAR(betas,1)$b
T <- length(ti)
q <- length(betas)
#priors (beta,vt|D0)
nu0 = 100; S0 = 0.3^2
m0 = c(1,1,0)
C0 = diag(diag(cov(betas)),nrow = 3)
#discount factor
b = 0.80; d = 0.95
sm = sa = matrix(rep(0,q*T),nrow = q); sC = sR = array(rep(0,q*q*T),dim = c(q,q,T)) 
sf = sQ = list()
sS = snu = rep(0,T)
nu <- nu0; S = S0; m <- m0; C <- C0
#Foreward filtering
for(t in 1:T){
  idx = data$date == ti[t]
  yt = data$lprice[idx]; ot = length(yt)
  X = as.matrix(cbind.data.frame(intercept = rep(1,sum(idx)),data[idx,-c(1,4)]))    #regression matrix
  
  #evolve
  nu <- b*nu
  sa[,t] <- a <- G%*%m
  sR[,,t] <- R <- G%*%C%*%t(G)/d

  #forecast
  sf[[t]] <- f <- X%*%a
  sQ[[t]] <- Q <- X%*%R%*%t(X) + diag(rep(S,ot),nrow = ot)
  
  #update
  Tt <- solve(t(X)%*%X + c(S)*solve(R))
  sm[,t] <- m <- Tt%*%(t(X)%*%yt + c(S)*solve(R)%*%a)
  sS[t] <- S <- (nu*c(S) + t(yt)%*%yt + c(S)*t(a)%*%solve(R)%*%a - t(m)%*%solve(Tt)%*%m)/(nu +ot)
  snu[t] <- nu <- nu + ot
  sC[,,t] <- C <- Tt/c(S)
}

```


##

FFBS with $\beta \sim VAR(1)$

```{r}
G <- VAR(betas,2)$b #Evolution matrix
G1 <- G[,1:4];G2 <- G[,5:8]
#G <- diag(rep(1.2,4),nrow = 4)
T <- length(ti)
q <- ncol(betas)
#priors (beta,vt|D0)
nu0 = 100; S0 = 0.3^2
m0 = c(1,1,0,0)
C0 = diag(diag(cov(betas)),nrow = 4)
#discount factor
b = 0.99; d = 0.99
sm = sa = matrix(rep(0,q*(T+2)),nrow = q); sC = sR = array(rep(0,q*q*(T+2)),dim = c(q,q,(T+2))) 
sf = sQ = list()
sS = snu = rep(0,T)
nu <- nu0; S = S0; m <- m0; C <- C0
sm[,1:2] <- rep(m0,2)
sC[,,1:2] <- rep(C0,2)
#Foreward filtering
for(t in 1:T){
  idx = data$date == ti[t]
  yt = data$lprice[idx]; ot = length(yt)
  X = as.matrix(cbind.data.frame(intercept = rep(1,sum(idx)),data[idx,-c(1,5)]))    #regression matrix
  
  #evolve
  nu <- b*nu
  sa[,t] <- a <- G1%*%sm[,t+1] + G2%*%sm[,t]
  sR[,,t] <- R <- (G1%*%sC[,,t+1]%*%t(G1) + G2%*%sC[,,t]%*%t(G2))/d

  #forecast
  sf[[t]] <- f <- X%*%a
  sQ[[t]] <- Q <- X%*%R%*%t(X) + diag(rep(S,ot),nrow = ot)
  
  #update
  Tt <- inv(t(X)%*%X + c(S)*inv(R))
  sm[,t+2] <- m <- Tt%*%(t(X)%*%yt + c(S)*inv(R)%*%a)
  sS[t] <- S <- (nu*c(S) + t(yt)%*%yt + c(S)*t(a)%*%inv(R)%*%a - t(m)%*%inv(Tt)%*%m)/(nu +ot)
  snu[t] <- nu <- nu + ot
  sC[,,t+2] <- C <- Tt/c(S)
}
sm <- sm[,-(1:2)]
sa <- sa[,-(1:2)]
#backward smoothing
mf <- sm; Cf <- sC; sf <- sS; nuf <- snu;
for(t in (T-1):1){
  sm[,t] <- (1-d)*sm[,t] + d*sm[,(t+1)]
  #if(t > p){
  #  e[t] <- x[t] - t(m[,t]) %*% x[(t-1):(t-p)]
  #}
  snu[t] <- (1-b)*snu[t]+ b*snu[t+1]
  st <- sS[t]; sS[t] <- 1/((1-b)/st + b/sS[t+1])
  sC[,,t] <- sS[t] * ((1-d)*sC[,,t]/st + d*d*sC[,,(t+1)]/sS[t+1])
}

```


```{r}
for(t in 1:T){
  idx = data$date == ti[t]
  yt = data$lprice[idx]; ot = length(yt)
  X = as.matrix(cbind.data.frame(intercept = rep(1,sum(idx)),data[idx,-c(1,5)]))    #regression matrix
  
  #evolve
  nu <- b*nu
  sa[,t] <- a <- G%*%m
  sR[,,t] <- R <- G%*%C%*%t(G)/d

  #forecast
  sf[[t]] <- f <- X%*%a
  sQ[[t]] <- Q <- diag(X%*%R%*%t(X) + diag(rep(S,ot),nrow = ot))
  
  #update
  e <- yt - f;
  A <- R%*%t(X)/matrix(rep(Q,q),nrow = q, byrow = T)
  
  r <- (nu+sum(e^2/q))/(nu + ot)
  snu[t] <- nu <- nu + ot
  sS[t] <- S <- S*r
  sm[,t] <- m <- a + A%*%e
  sC[,,t] <- C <- r*(R - (matrix(rep(Q,q),nrow = q, byrow = T)*A)%*%t(A))
```

```{r v1}
data <- apartment %>% 
  dplyr::select(lprice,lsize,lage,lfloor,date)
predictors <- c("lsize","lage","lfloor")
beta <- Tbeta(data,predictors)$beta
sigma <- Tbeta(data,predictors)$sigma
del <- c(0.95,0.95)
p <- 1
m0 <- c(1,1,1,0); C0 <- diag(diag(cov(beta)),nrow = ncol(beta));
n0 <- 100; S0 <- 0.3^2
lmvar <- function(data,beta,p,del,m0,C0,n0,S0){
  #dimension of process
  ti <- sort(unique(data$date))
  T <- dim(beta)[1]
  q <- dim(beta)[2]
  
  #evolution matrix based on beta
  Gs <- VAR(beta,p,"c")$b
  #trend <- Gs[,1]
  const <- Gs[,1]
  G <- array(Gs[,-1],dim = c(q,q,p))
  
  #return object allocation
  m <- a <- matrix(rep(0,q*(T+p)),nrow = q); C <- R <- array(rep(0,q*q*(T+p)),dim = c(q,q,(T+p))) 
  f <- Q <- list()
  S <- n <- rep(0,T)
  
  #initial values
  b <- del[1]; d <- eigen(Gs)$values[which.min(eigen(Gs)$values)] - 0.01
  St <- S0; nt <- n0
  m[,(1:p)] <- rep(m0,p); mt <- m0
  C[,,(1:p)] <- rep(C0,p); Ct <- C0
  
  #forwarding filtering
  for(t in 1:T){
    idx = data$date == ti[t]
    yt = data$lprice[idx]; ot = length(yt)
    X = as.matrix(cbind(rep(1,sum(idx)),data[idx,colnames(beta)[-1]]))    #regression matrix
  
    #evolve
    nt <- b*nt
    a[,t] <- at <- evolve(G=G,p=p,m = m[,(t+p-1):t])
    R[,,t] <- Rt <- evolve(G=G,p=p,C = C[,,(t+p-1):t])/d

    #forecast
    f[[t]] <- ft <- X%*%at
    Q[[t]] <- Qt <- X%*%Rt%*%t(X) + diag(rep(St,ot),nrow = ot)
  
    #update
    Tt <- solve(t(X)%*%X + c(St)*solve(Rt))
    m[,t+p] <- mt <- Tt%*%(t(X)%*%yt + c(St)*solve(Rt)%*%at)
    S[t] <- St <- (nt*c(St) + t(yt)%*%yt + c(St)*t(at)%*%solve(Rt)%*%at - t(mt)%*%solve(Tt)%*%mt)/(nt +ot)
    n[t] <- nt <- nt + ot
    C[,,t+p] <- Ct <- Tt/c(St) 
  }
  #backward smoothing
  mf <- m; Cf <- C; Sf <- S; nf <- n;
  #Gb <- rbind(Gs,diag(rep(1,dim(Gs)[2])))[1:dim(Gs)[2],]
  #mb <- rbind(mb[,T+p])
  Ct <- C[,,T]; St <- S[T]; nt <- n[T]
  B <- d*solve(G[,,1])
  #for(t in (T-1):1){
  #  m[,t] <- (1-d)*m[,t] + d*m[,(t+1)]
  #  n[t] <- (1-b)*n[t]+ b*n[t+1]
  #  St <- S[t]; S[t] <- 1/((1-b)/St + b/S[t+1])
  #  C[,,t] <- S[t]*((1-d)*C[,,t]/St + d*d*C[,,(t+1)]/S[t+1])
  #}
  for(t in (T-1):1){
    m[,t] <- (1-d)*m[,t] + B%*%(m[,(t+1)])
    n[t] <- (1-b)*n[t] + b*n[t+1]
    St <- S[t]; S[t] <- 1/((1-b)/St + b/S[t+1])
    C[,,t] <- S[t]*((1-d)*C[,,t]/St + B%*%C[,,(t+1)]%*%t(B)/S[t+1])
  }
}
```


```{r}
data <- apartment %>% 
  dplyr::select(lprice,lsize,lfloor,lage,date)
predictors <- c("lsize", "lfloor","lage")
beta <- Tbeta(data,predictors)$beta
del <- c(0.8,0.80)
p <- 1
m0 <- c(1,1,0,0); C0 <- diag(diag(cov(beta)),nrow = ncol(beta));
n0 <- 100; S0 <- 0.3^2
lmvar <- function(data,beta,p,del,m0,C0,n0,S0){
  #dimension of process
  ti <- sort(unique(data$date))
  T <- dim(beta)[1]
  q <- dim(beta)[2]
  
  #evolution matrix based on beta
  Gs <- VAR(beta,p)$b
  G <- array(Gs,dim = c(q,q,p))
  
  #return object allocation
  m <- a <- matrix(rep(0,q*(T+p)),nrow = q); C <- R <- array(rep(0,q*q*(T+p)),dim = c(q,q,(T+p))) 
  f <- Q <- list()
  S <- n <- rep(0,T)
  
  #initial values
  b <- del[1]; d <- del[2]
  St <- S0; nt <- n0
  m[,(1:p)] <- rep(m0,p); mt <- m0
  C[,,(1:p)] <- rep(C0,p); Ct <- C0
  
  #forwarding filtering
  for(t in 1:T){
    idx = data$date == ti[t]
    yt = data$lprice[idx]; ot = length(yt)
    X = as.matrix(cbind(rep(1,sum(idx)),data[idx,colnames(beta)[-1]]))    #regression matrix
  
    #evolve
    nt <- b*nt
    a[,t] <- at <- evolve(G=G,p=p,m = m[,(t+p-1):t])
    R[,,t] <- Rt <- evolve(G=G,p=p,C = C[,,(t+p-1):t])/d

    #forecast
    f[[t]] <- ft <- X%*%at
    Q[[t]] <- Qt <- X%*%Rt%*%t(X) + diag(rep(St,ot),nrow = ot)
  
    #update
    Tt <- solve(t(X)%*%X + c(St)*solve(Rt))
    m[,t+p] <- mt <- Tt%*%(t(X)%*%yt + c(St)*solve(Rt)%*%at)
    S[t] <- St <- (nt*c(St) + t(yt)%*%yt + c(St)*t(at)%*%solve(Rt)%*%at - t(mt)%*%solve(Tt)%*%mt)/(nt +ot)
    n[t] <- nt <- nt + ot
    C[,,t+p] <- Ct <- Tt/c(St) 
  }
  #backward smoothing
  mf <- m; Cf <- C; Sf <- S; nf <- n;
  #Gb <- rbind(Gs,diag(rep(1,dim(Gs)[2])))[1:dim(Gs)[2],]
  #mb <- rbind(mb[,T+p])
  Ct <- C[,,T]; St <- S[T]; nt <- n[T]
  G <- VAR(beta,1)$b
  B <- d*solve(G)
  for(t in (T-1):1){
    m[,t] <- (1-d)*m[,t] + B%*%m[,(t+1)]
    n[t] <- (1-b)*n[t]+ b*n[t+1]
    St <- S[t]; S[t] <- 1/((1-b)/St + b/S[t+1])
    C[,,t] <- S[t]*((1-d)*C[,,t]/St + B%*%C[,,(t+1)]%*%t(B)/S[(t+1)])
  }
}
```

