---
title: 'STA 532 Homework1'
author: "Jae Hyun Lee, jl914"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, message=F, warning=F, echo=F}
library(tidyverse)
require(magrittr)
require(plyr)
ggplot2::theme_set(ggplot2::theme_bw())
knitr::opts_chunk$set(fig.align = 'center')
```

# HW1 for STA-532


## Exercise 1

Let assume $\mathbb{Y} = \mathbb{R}$

#### (a)


$$
\begin{aligned}
Pr(Y \in (a,b]) = Pr(a<Y\le b) &= Pr(Y \le b) - Pr(Y \le a) \\
&= F(b) - F(a)
\end{aligned}
$$

#### (b)


$$
\begin{aligned}
Pr(Y \in (a,b)) = Pr(a<Y<b) &= Pr(Y < b) - Pr(Y \le a) \\
&= Pr(Y \le b ) - Pr(Y = b) - Pr(Y \le a) \\
&= F(b) - F(a) - Pr(Y = b)
\end{aligned}
$$

#### (c)


$$
\begin{aligned}
Pr(Y \in [a,b]) = Pr(a \le Y \le b) &= Pr(Y \le b) - Pr(Y < a) \\
&= Pr(Y \le b ) - (Pr(Y \le a) - Pr(Y = a)) \\
&= F(b) - F(a) + Pr(Y = a)
\end{aligned}
$$

## Exercise 2

#### (a)

$$
\begin{aligned}
&y \in (0,1) \rightarrow w = -log(y) \in (0,\infty) \\
&g(y) = -log(y) \; is \; monotonic \; and \; g^{-1}(w) = e^{-w} \\
\end{aligned}
$$


$$
\begin{aligned}
\rightarrow P_W(w) &= P_Y(g^{-1}(w))\mid\frac{d}{dw}g^{-1}(w)\mid \\
&= 1 \times e^{-w} = e^{-w} \; where \; w \in (0,\infty) \\
\\
&\int e^{-w}dw = 1 \rightarrow valid \; pdf
\end{aligned}
$$

#### (b)

$$
\begin{aligned}
&y \in \mathbb{R} \rightarrow w = y^{-1} \in \mathbb{R} \\
&g(y) = 1/y \; is \; monotonic \rightarrow g^{-1}(w) = 1/w
\end{aligned}
$$

$$
\begin{aligned}
\rightarrow P_W(w) &= P_Y(1/w) \mid \frac{d}{dw} \frac{1}{w} \mid \\
&= 1/(\pi(1+w^{-2})) \times 1/w^{-2} = \frac{1}{\pi(1+w^2)} \\
\\
&\int P_W(w)dw = \int P_Y(y)dy = 1
\end{aligned}
$$

#### (c)

$$
\begin{aligned}
&y \in \mathbb{R} \rightarrow w = e^y \in (0,\infty) \\
&P_Y(y) = \frac{1}{\sqrt(2\pi)} \\
&g(y) = e^y \; is \; monotonic \rightarrow g^{-1}(w) = log(w)
\end{aligned}
$$

$$
\begin{aligned}
\rightarrow P_W(w) &= P_Y(log(w)) \mid \frac{d}{dw} log(w) \mid \\
&= \frac{1}{\sqrt(2\pi)} w^{-1} exp\{-\frac{1}{2}(log(w))^2 \}
\end{aligned}
$$

#### (d)

$$
\begin{aligned}
&y \in \mathbb{R} \rightarrow w = y^2 \in [0,\infty) \\
&P_Y(y) = \frac{\Gamma((\nu+1)/2)}{\sqrt(\nu\pi) \Gamma(\nu/2)} (1+\frac{y^2}{\nu})^{-(v+1)/2} \\
&g(y) = y^2 \; is \; not \; monotonic
\end{aligned}
$$

$$
\begin{aligned}
\rightarrow F_W(w) &= Pr(W \le w) \\
&= Pr(Y^2 \le w) \\
&= Pr(-\sqrt{w} \le Y \le \sqrt{w} ) \\
\rightarrow P_W(w)& = [P_Y(\sqrt{w}) + P_Y(\sqrt{-w}))] \times 1/2\sqrt{w} \\
&= \frac{\Gamma((\nu+1)/2)}{\sqrt(\nu\pi) \Gamma(\nu/2)} (1+\frac{w}{\nu})^{-(v+1)/2} \times w^{-\frac{1}{2}}
\end{aligned}
$$

## Exercise 3

$0 = F(-\infty) < F(Y) < F(Y+\lambda) < F(\infty) = 1 \; for \; \lambda>0$
which means strictly increasing CDF and $F_Y^{-1}$ exists.

#### (a)

$$
\begin{aligned}
F_U(u) &= Pr(U \le u) \\
&= Pr(F_Y(Y) \le u) \\
&= Pr(Y \le F^{-1}_Y(u)) \\
&= F_Y(F_Y^{-1}(u)) = u \quad where\; u \in[0,1] \\
\\
valid \; check: &\int u du = 1
\end{aligned}
$$

#### (b)

$$
\begin{aligned}
F_X(x) &= Pr(X \le x) \\
&= Pr(F^{-1}_Y(U) \le x) \\
&= Pr(U \le F_Y(x)) \\
&= F_U(F_Y(x)) = F_Y(x) \\
\end{aligned}
$$
By Thm, X and Y are equivalent in distribution.

#### (c)

```{r}
Y = qnorm(runif(1000))
library(ggplot2)
ggplot() + geom_density(aes(x = Y))
```

## Exercise 4

Define $Pr(X \in A \mid Y = y)$

$$
\begin{aligned}
Pr(X \in A \mid Y = y) &= \frac{P(\{w:X(w) \in A, Y(w) = y\})}{P(\{w:Y(w) = y\})} \\
&= \frac{P(\{w:X(w) \in A, Y(w) = y\})}{P(\{w:X \in \mathbb{R},Y(w) = y\})} \\
&= \frac{\int_A P(x,y)dx}{\int_\mathbb{R}P(x,y)dx} \\
&= \frac{\int_A P(x,y)dx}{P(y)} = \int_A \frac{P(x,y)}{P(y)} dx = \int_A P_{X\mid Y}(x \mid y)dx
\end{aligned}
$$

Show that $Pr(X \in A \mid Y = y)P_Y(y)dy = Pr(X\in A)$

$$
\begin{aligned}
\int Pr(X \in A \mid Y = y)P_Y(y)dy &= \int \int_A P_{X|Y}(x\mid y)dx P_Y(y)dy \\
&= \int\int_A P_{X|Y}(x\mid y)P_Y(y)dxdy \\
&= \int_A\int P_{X|Y}(x\mid y)P_Y(y)dydx \quad (by \; Fubini's \; theorem) \\
&= \int_A\int \frac{P(x,y)}{P_Y(y)} P_Y(y)dydx \\
&= \int_A\int P(x,y)dydx \\
&= \int_A P_X(x)dx = Pr(X \in A)
\end{aligned}
$$

Since we integrate(sum) all probability $X\in A$ given Y = y over y,
It returns the probability that $X \in A$.
Assume that there are 5 people and a cup of water exists. After 5 minutes, probability that cup is empty is sum of probability that one of people drank that water.
In mathematical explanation, Let $\mathbb{Y} = \{1,2,\cdots,5\}$ and $P(\{w:Y(w) = i\}) = \frac{1}{5}$.

$$
\begin{aligned}
Pr(X \in A) &= P(\{w:X(w) \in A\}) \\
&=P(\cup_{i=1}^5\{w:X(w)\in A, y(w) = i\}) \\
&= \sum_{i=1}^5 P(\{w:X(w) \in A, y(w) = i\}) \\
&= \sum_{i=1}^5 P(\{w:X(w) \in A\})/5 = P(\{w:X(w) \in A\}) \\
\end{aligned}
$$

## Exercise 5

$B_{\epsilon} = (y - \epsilon,y]$ and $\lim_{\epsilon \to 0} B_\epsilon = \{y\}$

$$
\begin{aligned}
\lim_{\epsilon \to 0} Pr(X \in A \mid Y \in B_\epsilon) &= Pr(X\in A \mid Y = y)\\
&= \frac{P(X\in A, Y = y)}{P(Y = y)} \\
&= \frac{P(X \in A, Y = y)}{P_Y(y)} \\
&= \int_A \frac{P(x,y)}{P_Y(y)} = \int_A P_{X\mid Y}(x\mid y)dx
\end{aligned}
$$

## Exercise 6

$$
\begin{aligned}
P(x) = b^ax^{a-1}e^{-bx}/\Gamma(a) &\rightarrow \int^{\infty}_0 P(x)dx = \frac{b^a}{\Gamma(a)}\int x^{a-1}e^{-bx}dx = 1 \\
&\rightarrow \int x^{a-1}e^{-bx}dx = \frac{\Gamma(a)}{b^a} \\
\\
P_{Y\mid X}(y\mid x) = x^cy^{c-1}e^{-xy}/\Gamma(c)
\end{aligned}
$$

$$
\begin{aligned}
P_Y(y) &= \int P(x,y)dx \\
&= \int P(y\mid x)P(x)dx \\
&= \int x^cy^{c-1}e^{-xy}/\Gamma(c) \times b^ax^{a-1}e^{-bx}/\Gamma(a) dx \\
&= \frac{b^ay^{c-1}}{\Gamma(a)\Gamma(c)} \int x^{a+c-1}e^{-x(b+y)} dx \\
&= \frac{b^ay^{c-1}}{\Gamma(a)\Gamma(c)} \times \frac{\Gamma(a+c)}{(b+y)^{a+c}} \\
&= \frac{\Gamma(a+c)}{\Gamma(a)\Gamma(c)} \times (\frac{b}{b+y})^a \times (\frac{y}{b+y})^{c-1} \times \frac{1}{b+y}
\end{aligned}
$$

$$
\begin{aligned}
P_{X\mid Y}(x \mid y) &= \frac{P(x,y)}{P_Y(y)} \\
&= \frac{P_{Y \mid X}(y \mid x)P(x)}{P_Y(y)} \\
&= \frac{\frac{1}{\Gamma(a)\Gamma(c)} \times x^{a+c-1}e^{-x(b+y)}}{\frac{\Gamma(a+c)}{\Gamma(a)\Gamma(c)} \times (\frac{b}{b+y})^a \times (\frac{y}{b+y})^{c-1} \times \frac{1}{b+y}} \\
&= \frac{1}{\Gamma(a+c)} \times \frac{1}{b+y} \times (\frac{bx}{b+y})^a \times (\frac{yx}{b+y})^{c-1} \times e^{-x(b+y)}
\end{aligned}
$$